Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-13 paper/sheaf_lics.tex
add trunc

\usepackage{enumerate}
\usepackage{wasysym}
\usepackage[all]{xy}
\def\dar[#1]{\ar@<2pt>[#1]\ar@<-2pt>[#1]} %double arrows in xy
\def\tar[#1]{\ar@<2pt>[#1]\ar@<0pt>[#1]\ar@<-2pt>[#1]} %triple arrows in xy
\DeclareMathOperator{\Type}{Type}
\DeclareMathOperator{\HProp}{HProp}
\DeclareMathOperator{\nat}{nat}
\DeclareMathOperator{\Unit}{Unit}
\DeclareMathOperator{\im}{Im}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Contr}{Contr}
\newcommand{\modal}{\ensuremath{\ocircle}}

@@

We refer to~\cite{hottbook}.
Truncation levels + Im ?

If $x$ is an object of a product or a sigma type, we denote the $i$-th
projection of $x$ by either $\pi_i x$ or $x_i$.

\begin{itemize}
\item

We use the subobjects classification seen in~\cite{sets_in_hott}~: for
all type $B$, $\Type_n$ classifies subobjects of $B$ with $n$-truncated
fibers ; we have an equivalence
$$\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
\mathrm{IsTrunc}~n~\mathrm{fib}_f(b) \simeq (B \to \Type_n)$$
and a pullback (with $\Type_n^\bullet$ the universe of pointed
$n$-truncated types)~:
$$
\xymatrix{
  A \ar[r]^{t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\mathrm{pr}_1}\\
  B \ar[r]_{\chi_f} & \Type_n
}
$$
forall $f$ with $n$-truncated fibers, with 
$$t_f = \lambda a,~(\mathrm{fib}_f(f(a)),(a,\mathrm{idpath})).$$
\end{itemize}

@@

Since basic operations (dependent products, products, sigma types) are
stable for truncation levels, all theorem in~\cite{hottbook}, chapter
7.7 remains true.

@@

Another useful properties of modalities is the following~:
\begin{prop}
  If $P:\HProp$, then $\modal \widehat P : \HProp$, where $\widehat P$
  is $P$ seen as a $p$-type.
\end{prop}

A left exact modality leads to a new theory with new principles. A
basic example is the {\em open modality of $P$} for a $P:\HProp$, $\modal_P T = P \to T$. 
When one works in the subuniverse defined by this modality, defining
an inhabitant of a type $T$ is actually an inhabitant of $P \to
T$. Thus, being in this subuniverse is just adding the axiom $P$ to
the theory. Sadly, adding an axiom using the modality does not add any
computational content to $P$, thus it is not completely satisfactory.

Another example is the $\lnot\lnot$-modality, $\modal_{\lnot\lnot} T = (T
\to \bot) \to \bot$. Although we might think it enables us to use
classical reasonings in the corresponding reflective subuniverse, the
main problem is that every type in the new universe is an $\HProp$,
which is very restrictive.

@@

The Giraud axiom asserts that, if $f:X\to Y$ is a surjection, then the
colimit of $C(f)$ is $Y$. Viewing homotopy type theory as a
type-theoristic version of higher topos theory, it seems that this
axiom can be proved~\cite{lurie}, although we admit it in this paper.

@@

We now need to show that $\square$ is universal. The following
lemma is central in the proof~:
\begin{lem}
  Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
  diagonal diagram
  $$\cdots \overline{\Delta_3} \underset{\longrightarrow}{\stackrel{\longrightarrow}{\longrightarrow}}
   \overline{\Delta_2}
\stackrel{\longrightarrow}{\longrightarrow} \overline{\Delta_1},$$
where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
\end{lem}

@@

  Let $k:\nat$, let's show that 
\begin{align*}
  &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
    \land \mu_T t_{k-1} \\
  = &\mu_T t_k) = \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
  \land t_{k-1} = t_k)
\end{align*}
By induction on $k$, and the preservation of products by $\modal$, it
suffices to show that for all $a,b:T$, $\modal (a=b) = (\mu_T a =
\mu_T b)$, which is true by univalence.
\end{proof}

>>>>

\usepackage{enumerate}
\usepackage{wasysym}
\usepackage[all]{xy}
\def\dar[#1]{\ar@<2pt>[#1]\ar@<-2pt>[#1]} %double arrows in xy
\def\tar[#1]{\ar@<2pt>[#1]\ar@<0pt>[#1]\ar@<-2pt>[#1]} %triple arrows in xy
\DeclareMathOperator{\Type}{Type}
\DeclareMathOperator{\HProp}{HProp}
\DeclareMathOperator{\IsHProp}{IsHProp}
\DeclareMathOperator{\nat}{nat}
\DeclareMathOperator{\Unit}{Unit}
\DeclareMathOperator{\im}{Im}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Contr}{Contr}
\DeclareMathOperator{\IsContr}{IsContr}
\newcommand{\IsType}[1]{\mathrm{Is-}#1\mathrm{-type}}
\newcommand{\modal}{\ensuremath{\ocircle}}

@@

Our work is mainly based on the stratification by $\Type_n$~:
\begin{defi}
  $\IsType n$ is defined by induction on $n\geqslant -2$~:
  \begin{itemize}
  \item $\IsType {(-2)} X$ if $X$ is a contractible type, {\em ie} $X$
    is pointed by $c:X$, and every other point in $X$ is connected to $c$.
  \item $\IsType {(n+1)} X = \prod_{x,y:X} \IsType n (x=y).$
  \end{itemize}
  Then, $\Type_n = \sum_{X:\Type} \IsType n X$.
\end{defi}
For a type $X$, to be in $\Type_n$ means that path spaces of $X$ are
trivial after $n+1$ iteration.
We define some shortcuts
\begin{itemize}
\item $\IsContr := \IsType {(-2)}$ and $\Contr := \Type_{-2}$
\item $\IsHProp := \IsType {(-1)}$ and $\HProp := \Type_{-1}$
\end{itemize}

Using higher inductive types, one can define for any type $X$ $\|X\| : \HProp$
generated by 
\begin{itemize}
\item a function $|\cdot|_X : X \to \|X\|$,
\item paths $x=y$ for all $x,y:\|X\|$.
\end{itemize}
$\|X\|$ is called the {\em (propositional) truncation of $X$}. The recursion principle asserts~:
\begin{lem}
  For any $A:\Type$ and $B:\HProp$, if $f:A \to B$ then there is an
  induced $g:\|A\|\to B$ such that $g(|a|)= f(a)$ for any $a:A$.
\end{lem}
We refer to~\cite{hottbook} for more details.

Working in $\Type_n$ allows to use the subobjects classification as in~\cite{sets_in_hott}~: for
all type $B$, $\Type_n$ classifies subobjects of $B$ with $n$-truncated
fibers ; we have an equivalence
$$\left(\sum_{A:\Type} \sum_{f:A \to B} \prod_{b\in B}
\IsType n~\mathrm{fib}_f(b)\right) \simeq (B \to \Type_n)$$
and a pullback (with $\Type_n^\bullet$ the universe of pointed
$n$-truncated types)~:
$$
\xymatrix{
  A \ar[r]^{t_f} \ar[d]_f & \Type_n^\bullet \ar[d]^{\pi_1}\\
  B \ar[r]_{\chi_f} & \Type_n
}
$$
forall $f$ with $n$-truncated fibers, with 
$$t_f = \lambda a,~(\mathrm{fib}_f(f(a)),(a,\mathrm{idpath})).$$

@@

Since basic operations (dependent products, products, sigma types) are
stable for truncation levels, all theorem in~\cite{hottbook}, chapter
7.7 remains true.
In particular, if $X:\Type_p$ verifies $P(X)$, then for any $x,y:X$, $P(x=y)$.

@@

As we want, a left exact modality preserves $\HProp$~:
\begin{prop}
  If $P:\HProp$, then $\modal \widehat P : \HProp$, where $\widehat P$
  is $P$ seen as a $p$-type.
\end{prop}

A left exact modality leads to a new theory with new principles. A
basic example is the {\em open modality of $P$} for a $P:\HProp$, $\modal_P T = P \to T$. 
When one works in the subuniverse defined by this modality, defining
an inhabitant of a type $T$ is actually giving an inhabitant of $P \to
T$. Thus, being in this subuniverse is just adding the axiom $P$ to
the theory. Sadly, adding an axiom this way does not add any
computational content to $P$, thus is not completely satisfactory.

Another example is the $\lnot\lnot$-modality, $\modal_{\lnot\lnot} T = (T
\to \bot) \to \bot$. Although we might think it enables us to use
classical reasonings in the corresponding reflective subuniverse, the
main problem is that every type in the new universe is an $\HProp$,
which is very restrictive. 
We will use this modality only in the $\HProp$ layer, and try to
extend it to every $\Type_n$ without collapsing every type.

@@

The Giraud axiom asserts that, if $f:X\to Y$ is a surjection, then the
colimit of $C(f)$ is $Y$. Viewing homotopy type theory as a
type-theoristic version of higher topos theory, \cite{lurie} suggests that this
axiom can be proved, although we admit it in this paper.

@@

We now need to show that $\square$ is universal. The following
lemma is central in the proof~:
\begin{lem}
  Let $T:\Type_{n+1}$. Then $\square T$ is the colimit of the closed
  diagonal diagram
  $$\xymatrix{ 
    \cdots \overline{\Delta_3} \tar[r] & \overline{\Delta_2} \dar[r] &
    \overline{\Delta_1}
  }$$
where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
\end{lem}

@@

  Let $k:\nat$, let's show that 
  \begin{align*}
    &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
      \land \mu_T t_{k-1} \mu_T t_k)\\
    &= \sum_{t:T^k} \modal (t_1 = t_2 \land \cdots
      \land t_{k-1} = t_k)
  \end{align*}
By induction on $k$, and the preservation of products by $\modal$, it
suffices to show that for all $a,b:T$, $\modal (a=b) = (\mu_T a =
\mu_T b)$, which is true by univalence.
\end{proof}
